From david.champredon@phac-aspc.gc.ca Wed Dec  6 09:18:57 2023
Date: Wed, 6 Dec 2023 14:18:54 +0000
From: "Champredon, David (PHAC/ASPC)" <david.champredon@phac-aspc.gc.ca>
To: David Earn <earn@math.mcmaster.ca>
Subject: Re: [23w5151] BIRS workshop: e-mail list, contributions to report

Hi David, 

I hope all is well with you. 

Here is my BIRS abstract.

Best,
dC

________________________________________________________________________________
From: BIRS Workshop <no-replies@workshops.birs.ca>
Sent: Monday, November 27, 2023 18:48
To: Champredon, David (PHAC/ASPC) <david.champredon@phac-aspc.gc.ca>
Subject: [23w5151] BIRS workshop: e-mail list, contributions to report  
From "David Earn" <earn@math.mcmaster.ca> to 23w5151@workshops.birs.ca on Mon,
27 Nov 2023 18:48:05 -0500:
----------------------------------------------------------------------


Dear Colleagues,

Thank you for participating in our CANMOD BIRS workshop from 12-17 Nov
2023.  We are writing with a few followup points:

No one objected to sharing their e-mail address with the full group, so a
full list of all participants with their e-mail address and affiliation is
attached.

We, the organizers, are required to write a 10-15 page report about the
workshop.  This report will be placed permanently on the BIRS web site.
BIRS workshop reports normally contain summaries of the presentations.
Since we did not request abstracts in advance, we don't have summaries to
work from.  Consequently, if you gave a presentation at the meeting, can
you please send us a brief summary that can be included in the report?
Please note that the report must be formatted in LaTeX, so please send
LaTeX-compatible text if possible.  We would appreciate receiving this
brief text within two weeks, i.e., by Monday 11 Dec 2023, if at all
possible.  Thanks in advance!

Another common component of BIRS workshop reports comprises testimonials
from participants about the workshop.  If you have comments that you would
be comfortable appearing in the report, please also send them to us in a
LaTeX-friendly format, preferably by 11 Dec 2023.  Again, thanks in
advance.

Finally, if you attended in person and you have not already done so,
please submit your travel claim to Mark Campbell at SFU.  You should have
received an e-mail message from Mark at the end of the workshop.

Thanks again for participating in what worked out to be a very enjoyable
and interesting meeting!

Best wishes,
David, Caroline and Irena

------------------------------------------

   David J.D. Earn, PhD, FCAHS (he/???)
   Professor of Mathematics and
     Faculty of Science Research Chair in Mathematical Epidemiology
   Department of Mathematics and Statistics
     and M. G. DeGroote Institute for Infectious Disease Research
   McMaster University
   1280 Main Street West
   Hamilton, Ontario, Canada L8S 4K1

   E-mail:   earn@math.mcmaster.ca
   Homepage:https://can01.safelinks.protection.outlook.com/?url=http%3A%2F%2Fdavidearn.mcma
ster.ca%2F&data=05%7C01%7Cdavid.champredon%40phac-aspc.gc.ca%7Cf891f1e79fac4ff
a4e0508dbefa352fc%7C42fd9015de4d4223a368baeacab48927%7C0%7C0%7C638367256982488
995%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haW
wiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=6P7XezxYus1o4DWptcxY0tNuGPPFWSRfLLxa1e4
lMe4%3D&reserved=0
   COVID-19:https://can01.safelinks.protection.outlook.com/?url=https%3A%2F%2Fmac-theobio.g
ithub.io%2Fcovid-19%2F&data=05%7C01%7Cdavid.champredon%40phac-aspc.gc.ca%7Cf89
1f1e79fac4ffa4e0508dbefa352fc%7C42fd9015de4d4223a368baeacab48927%7C0%7C0%7C638
367256982488995%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLC
JBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=rd727I%2B1aDi72wtQaAnMTRwn2
JVfI3jFs8jQWyFObkY%3D&reserved=0
   Office:   Hamilton Hall, Room 317
   Tel:      (905) 525-9140, x27245





    [ Part 2, "abstract-champredon.tex"  Application/X-TEX (Name: ]
    [ "abstract-champredon.tex") 2.5 KB. ]
    [ Unable to print this part. ]
