
Infectious disease modelling that aims to contribute to public health
decision-making has a long history, going back to Daniel Bernoulli’s
work on smallpox control in the 18th century.  In the early 20th
century, Ronald Ross developed a mathematical model to help identify
effective malaria control strategies, and Kermack and McKendrick
developed the foundations of modern mathematical epidemiology.  While
theoretical work on disease modelling continued, attention from
decision-makers and politicians was scarce until the 2001 Foot and
Mouth Disease epidemic in the UK, and the SARS epidemic in 2003.  Over
the last 20 years, the perception of mathematical modelling as a
valuable tool in the public health policy process has become more and
more common, especially in the context of influenza pandemic
preparedness, which set the stage for immediate, serious engagement
with modellers when the SARS-CoV-2 pandemic exploded in early 2020.

From the start of the pandemic, many disease modellers around the
world found themselves in constant demand from public health agencies
and policy-makers.  Recognizing the importance of this development,
NSERC invested \$10M with the aim of creating networks of disease
modellers, public health professionals, and policy-makers.  A quarter of
the investment (\$2.5M) was allocated to our
\href{https://canmod.net/}{CANMOD network} (the remaining 75\% was
allocated to four other networks).

CANMOD aims to increase Canada’s capacity for data-driven emerging
infectious disease modelling (EIDM) to directly support short, medium,
and long-term public health decisions. Our network comprises
collaborative teams of modellers, statisticians, epidemiologists,
public health decision-makers, and those implementing and delivering
interventions. The questions we are tackling are grounded in public
health needs and generated in partnership between research
investigators and knowledge users -- public health leaders, health
administrators and policy-makers. This collaborative research supports
data collection, curation and access, with the hope that it will
increase the speed with which critical information is made available.


While the network initially focussed on addressing questions of
immediate relevance in the shorter term, we are continuing this
approach with longer-term challenges in mind, building a trajectory of
enhancing modelling expertise and capacity in infectious disease in
Canada. As researchers have moved from the daily challenges of
decision-making during the COVID-19 pandemic to working on longer-term
policy questions, CANMOD is building and coordinating national
capacity in infectious disease modelling at the forefront of public
health. This capacity will position public health across Canada for
better control of any infectious disease, and will build better
preparedness and resilience in case of future pandemics. While CANMOD
is primarily concerned with Canadian pandemic preparedness and
infectious disease control, the research resulting from the network’s
activities will provide an international benefit since emergent
pathogens pose a universal, global threat.

We are providing extensive experiential training opportunities for
postdoctoral fellows (PDFs), graduate and undergraduate students at
the intersection of infectious disease modelling, public health policy
and decision making. CANMOD is committed to increasing equity,
diversity, and inclusion in the next generation of infectious disease
modellers. Our trainees are well-placed for quantitatively-oriented
careers in academia, industry and the public sector, both in Canada
and abroad.

It is our experience that when modellers and statisticians undertake
research driven by public health and policy-relevant questions, it
leads to new science and methodology. If these questions were easy to
answer, public health researchers and institutions would answer them
with existing tools. However, even with established approaches to
infectious disease modelling, there remains a need to adapt and design
novel approaches and tools to address new and nuanced scientific
questions; to integrate noisy, incomplete, and abundant data streams;
and to capture fundamental properties defining local context. 

The
COVID-19 pandemic in Canada has made clear the urgent and immediate
need for modelling of local context to inform decisions that are often
implemented in local jurisdictions, and across diverse epidemiologic
and health system contexts. Effective public health benefits from the
support of engaged modellers who understand the local data, local
epidemiological, socio-cultural, and health-system contexts, and who
are passionate about collaborating on public health research
problems. Our 44 co-applicants and dozens of collaborators have been
engaged in this kind of work since the beginning of the COVID-19
pandemic, and are ideally placed to ensure that the collaborations we
continue to build are effective. The enthusiasm of public health
institutions is clear from the level of engagement to date, the
regular use of our research results in decision-making, and from
numerous and enthusiastic letters we have received from a wide range
of public health institutions across Canada, spanning municipal,
regional, provincial and national jurisdictions. Some of the
high-priority scientific questions that have emerged through this
collaborative research and were discussed at the workshop
are summarized below.

CANMOD’s multi-disciplinary and multi-sectoral network is addressing
all infectious disease challenges with principles of equality,
diversity and inclusion through its recruitment, training, and
research, and through events like the November 2023 BIRS meeting that
this report summarizes.
